---
title: "Social Network Analysis <br/> Home Assignment 3"
author: 'Maria Osipova'
date: "due date - 04.06.2018 23:59"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---
# Network communities
Please send your reports to hse.ntwks@gmail.com with the subject of of the following structure:<br/>
*[MAGOLEGO SNA 2018] {LastName} {First Name} HW{Number}*

Late submission policy: -1 point per day

Use this file as a template for your report.\
Support your computations with figures and comments. Send ONLY .Rmd versions of your report.

```{r, echo=FALSE}
library(igraph)
```


## Yahoo Music Network

For this assignment, you have to load [the part of Yahoo Music Network](https://drive.google.com/file/d/0Bwd5a7NScgMtNElfOFdSbE1UYnM/view). Archive contains network in GML format and corresponding list of artists.

```{r}
g <- read.graph("music2K.gml", format = "gml")
artists <- scan("artists.txt", what="", sep="\n")
V(g)$label = artists
```

Edges in this network appear if enough number of users have given ratings to both music bands. Note, that edges are weighted with similarity of the ratings.


### 1. Preprocessing

This network needs to be preprocessed. Do the steps, described on seminar, to make it look properly. How many vertices and edges were deleted?

```{r}
raw_vertex <- vcount(g)
raw_edges <- ecount(g)

g <- delete.vertices(g, degree(g) == 0)
is.simple(simplify(g, remove.multiple=FALSE))

clean_vertex <- vcount(g)
clean_edges <- ecount(g)

dv <- raw_vertex - clean_vertex
de <- raw_edges - clean_edges

print(sprintf("Was deleted %d vertices and %d edges", dv, de))
```

### 2. Clustering

Define clusters for this networks using one of the algorithms described on lectures and seminars:
```{r}
c <- fastgreedy.community(g)
```

Compute density of obtained clusters:
```{r}
dens <- 1:length(c)
inner_con <-  1:length(c)
for (i in 1:length(c)) {
  temp_g <- induced.subgraph(g, which(c$membership == i))
  
  e1 <- inner_con[i]
  d1 <- dens[i]
  
  e1 <- ecount(temp_g)
  d1 <- graph.density(temp_g)

  print(sprintf("Cluster ¹ %d with %d edges has density = %.4f", i, e1, d1))
  
}

```

Compute the ratio of inner clusters connections to outer ones:
```{r}
for (i in 1:length(c)){
  temp_g = induced.subgraph(g, c$membership == i)
  all_edges = length(E(g))
  inner_edges = length(E(temp_g))
  outer_edges = length(E(induced.subgraph(g,!(c$membership == i))))
  btw_edges = all_edges - inner_edges - outer_edges
  print(sprintf("Ratio for cluster # %d = %.3f",i, inner_edges/btw_edges))
}
```


### 3. Visualization & interpretation

Visulize five of the most dense clusters. Use names of artists as node labels on the graph.
```{r}
par(mfrow=c(2,3))
top_dens <- which(dens %in% sort(dens, decreasing = TRUE)[1:5])
for (i in 1:5) {
top_vert <- which(c$membership %in% top_dens[i])
top_graph <- induced.subgraph(g, top_vert)
plot(top_graph, layout=layout.circle, main=sprintf("Top %d dense cluster", i))
}
```

<b> (Extra task) </b>Try to interpret (using Wikipedia or other resources) obtained clusters.

*Top-1 cluster:  Cluster mostly consists of USA artists which were popular in 80's-90's and played music in such genres as pop-dance and freestyle.*
*Top-2 cluster: Musicians in this cluster are mostly American gospel singers.*
*Top-3 cluster: This cluster consists of popular stand-up comedians.*
*Top-4 cluster: Musicians in this cluster play funk music.*
*Top-5 cluster: Cluster consists of North-American jazz musicians.*
